---
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Reproducbile Research: Peer Assessment 1

## Loading and preprocessing the data

The activity data is loaded from a CSV-formatted file: 'activity.csv', with the columns:

* steps
* date
* interval

Read the data, and change the date column to POSIXct date form. Make use of the lubridate and dplyr libraries in all the following.

```{r load_date,message=FALSE}
require(lubridate)
require(dplyr)
activity <- read.csv('activity.csv')
activity$date <- ymd(activity$date)
```

## What is mean total number of steps taken per day?

To calculate the mean and median number of steps per day, we group the activity data by date, and then summarize the grouped data with the sum function.

```{r steps_per_day}
by_day <- group_by(activity,date)
steps <- summarize(by_day,sum.steps=sum(steps))
#plot(steps,
#     main='Steps per Day',xlab='Date (2012)',ylab='Number of Steps')
hist(steps$sum.steps,breaks=11,
     main='Distribution of Steps per Day',xlab='Steps per Day')
```

Compute the mean number of steps per day,

```{r}
mean(steps$sum,na.rm=T)
```

and the median number of steps per day.

```{r}
median(steps$sum,na.rm=T)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
